Timing and Positioning: Insights from Steak-Related Product Pricing at Walmart and Loblaws

### 2024-12-04 paper/paper.qmd
Update paper.qmd

This paper investigates pricing dynamics in the retail grocery sector, with a specific focus on steak-related products sold by two major supermarket chains: Walmart and Loblaws. The analysis leverages a rich dataset encompassing current and historical prices, product descriptions, and the timing of price changes to uncover trends in pricing strategies and their impact on consumer behavior. Steak-related products, such as AAA Angus Beef Ribeye Steak, serve as the centerpiece of this investigation, offering insights into the pricing tactics employed by each retailer. For instance, significant price reductions, such as a drop from $20.00/kg to $15.00/kg, provide a lens through which market competition and consumer demand can be analyzed. By focusing on these high-value products, the study aims to reveal how pricing adjustments influence purchasing patterns and shape the competitive landscape in the grocery sector.

>>>>

This paper investigates pricing dynamics in the retail grocery sector, focusing on steak-related products sold by two major supermarket chains: Walmart and Loblaws. The analysis leverages a rich dataset encompassing current and historical prices, product descriptions, and the timing of price changes to uncover trends in pricing strategies and their impact on consumer behavior. Steak-related products, such as AAA Angus Beef Ribeye Steak, serve as the centerpiece of this investigation, offering insights into the pricing tactics employed by each retailer. For instance, significant price reductions, such as a drop from $20.00/kg to $15.00/kg, provide a lens through which market competition and consumer demand can be analyzed. By focusing on these high-value products, the study aims to reveal how pricing adjustments influence purchasing patterns and shape the competitive landscape in the grocery sector.

The study of pricing dynamics in the retail grocery sector is crucial for understanding market behavior and consumer responses. As noted in the literature, "customer turnover responds to pricing,"@tonsor2009consumer. indicating that price adjustments can significantly impact consumer loyalty and purchasing decisions.Additionally, research has shown that "product variety conceptualized within brands does have an impact on aggregate subcategory sales,"@tonsor2009consumer. suggesting that both pricing and product assortment strategies are vital in influencing consumer behavior.These insights underscore the importance of strategic pricing and product management in the competitive retail grocery landscape.
